Unions of Balls and Configuration Space

### 2014-08-22 Writing/Final/article.tex
experiments

\begin{abstract}
Covering both the configuration space, or \emph{c-space}, and represent the right topology is hard to achieve at the same time with traditional sampling-based motion planning algorithms, such as Probabilistic Roadmap and Rapidly-Exploring Random Tree. In this paper, we proposed a method that is able to cover near optimal path and study the shape and topology of configuration space. Our method samples a serious of balls to cover the space and use the dual shape of the union of balls to re-construct the space. We analysis the coverage of the space and show how these balls and their dual shape can be used to generate a graph that maintains the topology of \emph{c-space}.
\end{abstract}

\section{Introduction}
\indent\indent Planning a collision-free motion for movable objects could be very challenge, especially in high dimensional space, because of the lack of knowledge to understand the shape and topology of configuration space. Meanwhile, motion planning has extensive applicationsto areas like augmented reality \cite{aug}, computer-aided design \cite{design} and bioinformatics \cite{bioinformatics}.

\indent Although we have seen some study in low dimensional space trying to characterize the topology of \emph{c-space} \cite{ConCompNum} \cite{interval}, these methods suffer dimension curse in higher dimensional space.

\indent Sampling-based methods for motion planning appear to be very successful \cite{Book}. These methods are usually proved to be \emph{probabilistically complete}, thus guarantee to find a path if it exists. Medial Axis sampling \cite{MAPRM} \cite{UMAPRM} attracted researchers attention for their property of characterizing the topology of \emph{C-space} and safe clearance to obstacles, however, no structures are made with these samples to properly represent the space.

\indent Approximating the "shape" of a given point set in a space is a frequently occurred question in many research areas, which is the intuitive of $\alpha$-shape. A relative notion is space-filling diagram which is used in chemistry and biology to represent molecular structures. \cite{alpha intro}.

\indent Trying to answer the question "does a collision free path exist", Zoe McCarthy uses samples to build an $\alpha$-shape to better understand samples and space captured, which inspires us to explore more with $\alpha$-shape.

\indent In this paper, we will firstly introduce a sampling algorithm to cover the \emph{c-space}, medial axis sampling methods are applied to reduce the number of samples and capture the topology of the space. We then use these samples, or balls, to construct the weighted $\alpha$-shape. After simplifying the $\alpha$-shape, we will still have a simple graph, which can be used for both motion planning and topology study. In each step, we will demonstrate algorithm results of a 2D configuration space. Some properties of our algorithm are discussed in section 4.

@@

\indent\indent In this section, we first discuss the basics of sampling-based motion planning and the method of sampling on medial axis. With the information of positions and their clearance in c-space, we have samples as spheres, or balls. We then discuss the dual shape of the union of balls which plays an important role in understanding the shape and topology of \emph{c-space}. Some works on graph theory and topology will finally be introduced.s

@@

\paragraph{A} \emph{Sampling-based Motion Planning} \hfill \\
\indent A robot is a movable object whose state can be described by $n$ parameters, or \emph{degrees of freedom} ($DOFs$). A point $<x_1, x_2, ..., x_n>$ in a n-dimensional space uniquely defines the configuration of a robot. Such space is called \emph{"Configuration Space"} or \emph{"C-space"}  ($C$). The subset of all feasible (collision-free) configurations is called the \emph{free space} (\emph{$C_{free}$}), while $C_{obst} = C \setminus C_{free}$ is the union of all infeasible configurations. \cite{UMAPRM} Path planning will generally be viewed as a search in a metric space $X$ for continuous paths from an initial state $x_{init}$ to a goal region $X_{goal} \subset X$. For a standard problem, $X = C$. \cite{RRT} What needs to be pointed out is that \emph{C-space} is very different from workspace which is usually formed by polygons. The shape of \emph{C-space} is usually unknown to us, unless the robot is a point, in which case workspace is the \emph{C-space}.

\indent Sampling-based algorithms have been very successful and have seen many applications in industry for planning in high dimension space. Probablistic Roadmap \cite{PRM} and Rapidly-Exploring Random Tree \cite{RRT} are two outstanding methods. The basic idea under these methods is to randomly sample feasible configurations in \emph{C-space} and connect nearby valid samples as edges to construct a graph or tree structure, which careless about the dimension problem. By connecting start and goal configurations to the roadmap, a graph search, e.g., A*, can be performed to extract a near optimal solution path.

\indent One disadvantage of these methods is that we can't study the topological structure of C-space using only the data structures constructed.

%====================================
% Sampling on medial axis
%====================================
\paragraph{B} \emph{Medial Axis Sampling} \hfill \\
\indent A simpler method that maintains the topology of \emph{C-space} but doesn't cover near optimal path is medial axis sampling. Medial Axis is also called Generallized Voronoi Diagram. The medial axis $MA(C_{free})$ of free space $C_{free}$ is a strong deformation retract of $C_{free}$, which means $C_{free}$ can be continuously deformed onto $MA(C_{free})$ while maintaining its topological structure. However, medial axis is hard to compute explicitly.

\indent In Figure \cite{Book}, sec. 5.6.3, Steven LaValle defined Medial Axis as "Let $(X, \rho)$ be a metric space. Let a \emph{maximal ball} be a ball $B(x,r) \subseteq X$ such that no other ball can be a proper subset. The centers of all maximal balls trace out a one-dimensional set of points referred to as the \emph{medial axis}".

@@

\indent Steven A. Wilmart \cite{MAPRM} sampled on the medial axis of free space to build up a probabilistic roadmap planner. He also showed that "it is possible to efficiently retract a configuration, free or otherwise, onto the medial axis of the free space without having to compute the medial axis explicitly".

@@

\indent Hsin-Yi (Cindy) Yeh \cite{UMAPRM} solved this problem by putting some random "sticks" with random directions in free space. Any surfaces will intersect uniformly with these sticks, thus searching on sticks to find points that intersect with medial axis will give nodes uniformly distributed on the medial axis.

@@

\begin{figure}
	\center{\includegraphics[scale=0.35]{./pics/MA.png}}
	\caption{\label{fig:Medial Axis} Sampling on the Medial Axis of a rectangle space.\cite{Book}}
\end{figure}

@@

\indent Zoe McCarthy \cite{PathNonexistance} used samples generated by PRM to form the $\alpha$-shape of \emph{C-space}. She then proved there doesn't exist a path between a pair of configurations by proving they are in two different disconnected components of the shape. She declared it was the first time $\alpha$-shape is being used in motion planning.

\indent Define a \emph{simplicial complex} $D$: a collection of simplices such that if $\bigtriangleup_T \subseteq D$ ( $\bigtriangleup_T$ is the convex hull of $T$ ), then if $U \subseteq T$, we have $\bigtriangleup_U \subseteq D$ and also that the intersection of two simplices in $D$ is either empty or another simplex in $D$. The $\alpha$-shape of a set of points $S$ is a generalization of the convex hull of those points. \cite{PathNonexistance}. H. Edelsbrunner first introduced them in \cite{PointSetShape}. By choosing different values of the one parameter $\alpha$, we can have a family of shapes with the same set of points. An example of $\alpha$-shape is in Figure \ref{fig:Alpha-shape}.

@@

\begin{figure}
	\center{\includegraphics[scale=0.7]{./pics/alphaShape.png}}
	\caption{\label{fig:Alpha-shape} An example of $\alpha$-shape. \cite{PointSetShape}}
\end{figure}

\begin{figure}
	\center{\includegraphics[scale=0.30]{./pics/dualshape.png}}
	\caption{\label{fig:Dual Shape} An example of weighted $\alpha$-shape for a union of balls. \cite{DualShape}}
\end{figure}

@@

\indent The first betti number of a graph is very easy to compute. Given a graph $G$, with $n$ vertices, $m$ edges and $k$ components. It's 1st betti number is:

$\beta_1 = m - n + k$

\indent This can be proved straightforwardly by mathematical induction on the number of edges. A new edge either increments the number of 1-cycles or decrements the number of connected components. \cite{betti wiki}

\indent H. Edelsbrunner introduced a way to compute \emph{n-th} Betti number in \cite{Edel Book}, where he uses gaussian ellimination to compute the betti numbers with adjacency matrix.

\indent Vanessa Robins \cite{alpha betti} provided more detailed explains and proofs on computing betti numbers. What's more, she put more thoughts on computing betti numbers for alpha shapes.

\begin{figure}
	\center{\includegraphics[scale=0.26]{./pics/Torus.png}}
	\caption{\label{fig:Torus} The torus has one connected component($\beta_0$), two circular holes ($\beta_1$,the one in the center and the one in the middle of the "donut"), and one two-dimensional void ($\beta_2$, the inside of the "donut").\cite{betti wiki}}
\end{figure}

@@

\indent \indent Traditionally, sampling based algorithms will build a graph for configuration space. These kind of graphs are proved to be able to find near optimal path, but can hardly be used to analysis the topology of paths. We want to to build a stronger representation of \emph{C-space} that is both able to cover optimal path and study the topological structure. In this section, we are trying to solve the problem in 2D and answer the question "Is two paths in the same homotopy class" as a start point of using our method to study the topology of \emph{C-space}. In this section, we will first give the outline of the algorithm, and explain each step in detail.

%====================================
% Algorithm outline
%====================================
\paragraph{A} \emph{The Algorithm outline} \hfill \\
\indent Before starting introducing our algorithm, we firstly clarify some preliminary tools we can have: We are sampling in $(X, \rho)$ metric space. Let $x$ be a configuration in $X$(usually \emph{C-space}), $\rho(x)$ gives the clearance of configuration $x$. By \emph{"clearance"}, we mean its (estimated) distance to the nearest obstacles. Ball $B(x, \rho(x))$ is referred to as a $"sample"$ in the rest of the section. We make no assumption of the shape of obstacles regions in \emph{C-space}.

@@

%%%%%%%%%%%%%%%%================================================>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
%%%%%%  We need to name our algorithm
%%%%%%  Seriously
%---------------------------------------------------------------------------------------------
\begin{figure}
  \begin{algorithmic}[1]
  \Function{WeNeedAGoodName}{\emph{C-space}}

@@

  \EndFunction
  \end{algorithmic}
  \caption{\label{fig:Algo} Algorithm Outline}
\end{figure}

%====================================
% Covering c-space
%====================================
\paragraph{B} \emph{Sampling in C-space} \hfill \\
\indent As pointed out in Figure \ref{fig:Algo}, this phase has two parts: Sampling on medial axis, and sampling the rest of free space.

@@

\indent Getting configurations inside a ball doesn't have to be $O(n)$ costly. It can be done using KNN algorithm, which is more efficient than iterating everything. Since we generated random configurations $n$ times(not all are feasible), in a 2D space with area $A$. We expect to have $n/A$ configurations in an unit area. A ball with radius $r$ has area $\pi r^2$, approximately $k = \frac{n \pi r^2}{A}$ configurations are inside the ball. We then project these (uniformly distributed) configurations to the boundary of the ball to sample new spheres. Actually, choosing a larger $k$ could be better: we virtually extend the radius of our discs, which results in a smaller density in arcs that are very close to obstacles than those far from obstacles. Samples got in this part are called \emph{regular samples}.

@@

\indent In the first part, constructing the dual shape using only the medial axis sample could break the topology of medial axis. The reason is shown in Figure \ref{badthing}: three balls are all centered on medial axis, but their dual shape will be an empty triangle, which introduces a "hole" in that area. To solve this problem we can try to sample a ball centered at the the radical center after getting an empty triangle, and rebuild the dual shape for these four balls. \cite{radical center} Proving the solution is very simple: if there is no obstacles in that area, the newly sampled ball will cover the space, if there is obstacle, the ball will either not be sampled or not connect three of the balls.

@@

\indent The second part of this phase can be stated as: retract all connected 2-simplices into one point, reconnect these retracted points with other ones using their previous connection. By doing so, we got a simple graph that characterize the topology of \emph{C-space}. We would like to call it \emph{Topology Roadmap}. This roadmap can also be used to determine if paths exist between two configurations. An example of such roadmap is shown in Figure \ref{simplifiedgraph}

@@

\begin{figure}
	\center{\includegraphics[scale=0.3]{./pics/my_dualshape.PNG}}
	\caption{\label{ourdualshape}} Dual shape of balls generated in phase 1 (centered on medial axis).
\end{figure}

@@

\indent Determine if a simple graph has a loop is very easy using the first Betti number $\beta_1$ mentioned in section 2.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Properties of Our Algorithm
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Good Properties}\label{properties}

  \begin{theorem}
  Let $S_{init}$ be the set of initial random configurations. With minimum radius of $r_{min}$, configurations with clearance larger than $r_{min}$ will be covered, when $|S_{init}| \rightarrow \infty$.
  \end{theorem}

  \paragraph{\emph{Proof:}} 
  \indent When the number of initial random configurations approaches infinite, the boundaries of all existing balls are fully covered. So, every configuration not inside any existing balls will be sampled as a ball, unless it's less than $r_{min}$ clearance.

  \begin{theorem}
  Let $B$ be the set of ball samples got in phase one. Then $|B|$ is finite.   
  \end{theorem}
  \paragraph{\emph{Proof:}} 
  \indent We restrict balls to have no less than $r_{min}$ radius and new balls to be centered outside existing ones. Assume we partition the space into $r_{min} \times r_{min}$ cells in 2D. The number of cells is finite, if the space is finite. Putting balls with exactly $r_{min}$ radius in the centers of these cells will cover the space totally. The number of balls is finite. Phase one will generate balls no more than the number of balls generated in this way. Figure \ref{ourdualshape} shows the dual shape of balls generated by previous phase.

@@

  \begin{theorem}
  Any path $\sigma$ can be continuously deformed into medial axis. 
  \end{theorem}   
  \paragraph{\emph{Proof}}
  \indent This is the definition of Medial Axis $MA(C_{free})$ being a strong deformation retraction of $C_{free}$.

  \begin{theorem}
  Given path $\sigma_1$ and $\sigma_2$, we say they have the same homotopy type if their image in the topology roadmap is the same. 
  \end{theorem}
  \paragraph{\emph{Proof}}
  \indent Let $x$ be a point in the space $X$, $B = \{ b_1, b_2, ..., b_n \}$ be the set of balls centered on medial axis. Define map $H(x)$:

  $H(x) = b_i \indent, if \indent |x-b| \leq |x-b'|, b' \in B$

  \indent This continuous function maps any points in the space to balls centered on medial axis. While every point in the balls is mapped to the dual shape form by filled triangles, nodes and edges. Filled triangles are mapped to single nodes.

>>>>

\begin{abstract}
Covering both the configuration space, or \emph{C-space}, and represent the right topology is hard to achieve at the same time with traditional sampling-based motion planning algorithms, such as Probabilistic Roadmap and Rapidly-Exploring Random Tree. In this paper, we proposed a method that is able to cover near optimal path and study the shape and topology of configuration space. Our method samples a serious of balls to cover the space and use the dual shape of the union of balls to re-construct the space. The output of our algorithm can be viewed as a rich representation of \emph{C-space}. Benefits include (1)low memory for planning, (2) ability of analyzing shape and topology of \emph{C-space} and (3) fast planning for large space. We analysis the coverage of the space and show how these balls and their dual shape can be used to generate a graph that maintains the topology of \emph{C-space}.
\end{abstract}

\section{Introduction}

\indent\indent A robot state can be represented using n parameters $<x_1, x_2, ..., x_n>$ as a point in a n-dimensional space $X$, called \emph{Configuration Space}, or \emph{C-space}\cite{UMAPRM}. In this paper, we present a new method to model \emph{C-space}. Our algorithm is sampling-based, We will sample balls in \emph{C-space} to cover the space. By using the dual shape of unions of balls we construct the shape of free space, and capture its topology. A topology roadmap will then be built which we can use to determine path homotopy type, to find paths in different homotopy classes, and to answer other topology questions. We implemented the algorithm in a 2D world with a point robot.

\indent Representing \emph{C-space} is very fundamental for motion planning. Covering the space and characterize the topology are two concerns. By covering the space better, we can find better paths. By studying its topology, we can answer questions like "does a path exist between two configurations?", "does two paths has the same topology?" and "can we find two path that have different homotopy types?". These questions arise frequently in applications as predicting paths for unpredictable entities and deployment of multiple agents for efficient exploration of an environment\cite{pathtop}. Because of the extensive applications of motion planning to areas like augmented reality \cite{aug}, computer-aided design \cite{design} and bioinformatics \cite{bioinformatics}. Answering these questions could benefit many researches.

\indent Previous methods have their disadvantages. Some sampling-based methods, such as \emph{PRM*} and \emph{RRT*} \cite{prmstar}, are able to cover the space, but failed to capture the topology \cite{sampissues} because of ``holes" in the roadmap and inability to determine if a ``hole" contains obstacles. \emph{Visibility-base Probabilistic Roadmap}, or \emph{vis-PRM} \cite{vprm}, tried to capture the topology by placing ``guards" to ``see" the space. However, they can't describe ``visible regions" and overlapping regions explicitly, which makes it difficult to analyze coverage and find optimal paths. Medial Axis sampling attracted researchers attention, such as \emph{MAPRM}\cite{MAPRM} and \emph{UMAPRM}\cite{UMAPRM}, for their property of characterizing the topology of \emph{C-space} and safe clearance to obstacles, however, no structures are made with these samples to properly represent the space. We have seen some studies in low dimensional space trying to characterize the topology of \emph{C-space}, such as \cite{ConCompNum} and \cite{interval}, these methods suffer dimension curse in higher dimensional space. In \cite{pathtopsearch} and \cite{pathtop}, Bhattacharya, et al, studied searching path in a graph with topological constrains, and their method can be extended to higher dimensional space

%\indent Some sampling-based methods, such as \emph{PRM*} and \emph{RRT*} \cite{prmstar}, are able to cover the space, but failed to capture the topology \cite{sampissues} because of "holes" in the roadmap and inability to determine if a "hole" contains obstacles. Medial Axis sampling attracted researchers attention, such as \emph{MAPRM}\cite{MAPRM} and \emph{UMAPRM}\cite{UMAPRM}, for their property of characterizing the topology of \emph{C-space} and safe clearance to obstacles, however, no structures are made with these samples to properly represent the space. Another sampling-based algorithm that captures the topology of the space is \emph{Visibility-base Probabilistic Roadmap}, or \emph{vis-PRM} \cite{vprm}. Although \emph{vis-PRM} has some issues with complex narrow passages, using some sampling heuristics, like \emph{MAPRM}, such issues will be avoid.

\indent Our method addresses these problems by sampling weighted configurations, or balls, and constructing the dual shape of union of balls. Thus we are able to analyze the coverage of the space and answer topology-related questions. Besides, our algorithm has many advantages. It requires low memory for planning paths, allows analysis of shape and topology of \emph{C-space} and fastens planning for large space.

%\indent With sampling-based algorithms, planners will get a lot of feasible (collision-free) samples. These samples can actually be used to study the shape of \emph{C-space}. Approximating the "shape" of a given point set in a space is a frequently occurred question in many research areas, which is the intuitive of $\alpha$-shape. A relative notion is space-filling diagram which is used in chemistry and biology to represent molecular structures\cite{alpha intro}. McCarthy, et al, used this idea to answer the question "does a path exist between two configurations", which inspired us to explore more with $\alpha$-shape.

\indent In this paper, we are sampling in $(X, \rho)$ metric space. Let $x$ be a configuration in $X$(usually \emph{C-space}), $\rho(x)$ gives the clearance of configuration $x$, negative if infeasible, positive if feasible. By \emph{"clearance"}, we mean its (estimated) distance to the nearest obstacles. Our experiments environment is a point robot with polygonal obstacles, so we can fast determine the clearance of a configuration. ( In practice, workspace doesn't have to be the same as configuration space, as long as we can get the clearance of a configuration which can be done by checking the collision distance of a robot. )

%\indent\indent Planning a collision-free motion for movable objects could be very challenge, especially in high dimensional space, because of the lack of knowledge to understand the shape and topology of configuration space. Meanwhile, motion planning has extensive applicationsto areas like augmented reality \cite{aug}, computer-aided design \cite{design} and bioinformatics \cite{bioinformatics}.

%\indent Although we have seen some study in low dimensional space trying to characterize the topology of \emph{C-space} \cite{ConCompNum} \cite{interval}, these methods suffer dimension curse in higher dimensional space.

%\indent Sampling-based methods for motion planning appear to be very successful \cite{Book}. These methods are usually proved to be \emph{probabilistically complete}, thus guarantee to find a path if it exists. Medial Axis sampling \cite{MAPRM} \cite{UMAPRM} attracted researchers attention for their property of characterizing the topology of \emph{C-space} and safe clearance to obstacles, however, no structures are made with these samples to properly represent the space.

%\indent Approximating the "shape" of a given point set in a space is a frequently occurred question in many research areas, which is the intuitive of $\alpha$-shape. A relative notion is space-filling diagram which is used in chemistry and biology to represent molecular structures. \cite{alpha intro}.

%\indent Trying to answer the question "does a collision free path exist", Zoe McCarthy uses samples to build an $\alpha$-shape to better understand samples and space captured, which inspires us to explore more with $\alpha$-shape.

%\indent In this paper, we will firstly introduce a sampling algorithm to cover the \emph{C-space}, medial axis sampling methods are applied to reduce the number of samples and capture the topology of the space. We then use these samples, or balls, to construct the weighted $\alpha$-shape. After simplifying the $\alpha$-shape, we will still have a simple graph, which can be used for both motion planning and topology study. In each step, we will demonstrate algorithm results of a 2D configuration space. Some properties of our algorithm are discussed in section 4.

@@

\indent\indent In this section, we first discuss the basics of sampling-based motion planning and the method of sampling on medial axis. With the information of positions and their clearance in \emph{C-space}, we have samples as spheres, or balls. We then discuss the dual shape of the union of balls which plays an important role in understanding the shape and topology of \emph{C-space}. Some works on graph theory and topology will finally be introduced.s

@@

\paragraph{A} \emph{Sampling-based Motion Planning} \hfill \\
\indent A robot is a movable object whose state can be described by $n$ parameters, or \emph{degrees of freedom} ($DOFs$). A point $<x_1, x_2, ..., x_n>$ in a n-dimensional space uniquely defines the configuration of a robot. Such space is called \emph{"Configuration Space"} or \emph{"C-space"}  ($C$). The subset of all feasible (collision-free) configurations is called the \emph{free space} (\emph{$C_{free}$}), while $C_{obst} = C \setminus C_{free}$ is the union of all infeasible configurations\cite{UMAPRM}. Path planning will generally be viewed as a search in a metric space $X$ for continuous paths from an initial state $x_{init}$ to a goal region $X_{goal} \subset X$. For a standard problem, $X = C$ \cite{RRT}. What needs to be pointed out is that \emph{C-space} is very different from workspace which is usually formed by polygons. The shape of \emph{C-space} is usually unknown to us, unless the robot is a point, in which case workspace is the \emph{C-space}.

\indent Sampling-based algorithms have been very successful and have seen many applications in industry for planning in high dimension space. Probablistic Roadmap \cite{PRM} and Rapidly-Exploring Random Tree \cite{RRT} are two outstanding methods. The basic idea under these methods is to randomly sample feasible configurations in \emph{C-space} and connect nearby valid samples as edges to construct a graph or tree structure, which careless about the dimension problem. By connecting start and goal configurations to the roadmap, a graph search, e.g., A*, can be performed to extract a near optimal solution path. Problem is it's hard to analyze coverage of the space with these methods. And ``holes" in a roadmap cannot tell if there are obstacles, thus we can't use it to analysis path topology. Our method can avoid these problems by analyzing balls and their dual shape.

\indent Some methods do captures topology of \emph{C-space}. \emph{Visibility-based Probabilistic Roadmap}\cite{vprm} might be one of them. By randomly sampling ``guards" that can't see each other, the algorithm can produce guards that cover most of the space. Then connection nodes will be generated to connect guards. Our method is very much related to \emph{Vis-PRM}. ``Guards" of \emph{Vis-PRM} describe their ``visible regions" implicitly, while our method limit a configuration to have a ``visible range", the clearance. Thus, we can describe overlapping regions and visible regions explicitly.

\indent It's still possible to use \emph{PRM} to find paths with topology constrains. In \cite{pathtopsearch} and \cite{pathtop}, by combining integration theory with search techniques, Bhattacharya, et al, studied searching path in a graph with topological constrains, and their method can be extended to higher dimensional space.

\indent Methods like \emph{PRM*} requires much memory to store the roadmap because of the large amount of vertices and greater amount of edges. Some algorithms, such as \cite{weifu} and \cite{bekris}, have been proposed to reduce the memory requirement by reducing edges of \emph{PRM*}. These methods try to solve the problem after it happened, while our method tries to avoid the problem from happening.

%====================================
% Sampling on medial axis
%====================================
\paragraph{C} \emph{Medial Axis Sampling} \hfill \\
\indent A simpler method that maintains the topology of \emph{C-space} but doesn't cover near optimal path is medial axis sampling. Medial Axis is also called Generallized Voronoi Diagram. The medial axis $MA(C_{free})$ of free space $C_{free}$ is a strong deformation retract of $C_{free}$, which means $C_{free}$ can be continuously deformed onto $MA(C_{free})$ while maintaining its topological structure. However, medial axis is hard to compute explicitly.

\indent In \cite{Book}, sec. 5.6.3, S. LaValle defined Medial Axis as "Let $(X, \rho)$ be a metric space. Let a \emph{maximal ball} be a ball $B(x,r) \subseteq X$ such that no other ball can be a proper subset. The centers of all maximal balls trace out a one-dimensional set of points referred to as the \emph{medial axis}".

@@

\indent S. Wilmart \cite{MAPRM} sampled on the medial axis of free space to build up a probabilistic roadmap planner. He also showed that "it is possible to efficiently retract a configuration, free or otherwise, onto the medial axis of the free space without having to compute the medial axis explicitly".

@@

\indent Yeh \cite{UMAPRM} solved this problem by putting some random "sticks" with random directions in free space. Any surfaces will intersect uniformly with these sticks, thus searching on sticks to find points that intersect with medial axis will give nodes uniformly distributed on the medial axis.

@@

\begin{figure}
	\center{\includegraphics[scale=0.35]{./pics/MA.PNG}}
	\caption{\label{fig:Medial Axis} Sampling on the Medial Axis of a rectangle space. The black blocks are obstacles. We will use this world as our working space in the following sections. }
\end{figure}

@@

\indent We have seen papers concerning the shape of \emph{C-space}. Robin, et al, developed a computational method for detecting large convex regions of obstacle-free space\cite{largeconvex}. Knowing such information can be helpful to plan robot footstep locations while avoiding obstacles and other optimization problems. But this method can only capture local features.

\indent Z. McCarthy, et al, \cite{PathNonexistance} used samples generated by PRM to form the $\alpha$-shape of \emph{C-space}. She then proved there doesn't exist a path between a pair of configurations by proving they are in two different disconnected components of the shape. She declared it was the first time $\alpha$-shape is being used in motion planning.

\indent Define a \emph{simplicial complex} $D$: a collection of simplices such that if $\bigtriangleup_T \subseteq D$ ( $\bigtriangleup_T$ is the convex hull of $T$ ), then if $U \subseteq T$, we have $\bigtriangleup_U \subseteq D$ and also that the intersection of two simplices in $D$ is either empty or another simplex in $D$. The $\alpha$-shape of a set of points $S$ is a generalization of the convex hull of those points\cite{PathNonexistance}. H. Edelsbrunner first introduced them in \cite{PointSetShape}. By choosing different values of the one parameter $\alpha$, we can have a family of shapes with the same set of points. An example of $\alpha$-shape is in Figure \ref{fig:Alpha-shape}.

@@

\begin{figure}
	\center{\includegraphics[scale=0.7]{./pics/alphaShape.png}}
	\caption{\label{fig:Alpha-shape} An example of $\alpha$-shape}
\end{figure}

\begin{figure}
	\center{\includegraphics[scale=0.30]{./pics/dualshape.png}}
	\caption{\label{fig:Dual Shape} An example of weighted $\alpha$-shape for a union of balls. Copied from \cite{DualShape}}
\end{figure}

@@

\indent The first Betti number of a graph is very easy to compute. Given a graph $G$, with $n$ vertices, $m$ edges and $k$ components. It's 1st Betti number is

\begin{equation}
\beta_1 = m - n + k
\end{equation}

\indent Proof of this equation can be found in \cite{cyclomatic}.

\indent H. Edelsbrunner introduced a way to compute \emph{n-th} Betti number in \cite{Edel Book}, where he uses gaussian elimination to compute the Betti numbers with adjacency matrix.

\indent V. Robins \cite{alpha betti} provided a computational technique for computing Betti numbers with detailed explains and proofs. What's more, she put more thoughts on computing Betti numbers for alpha shapes.

\begin{figure}
	\center{\includegraphics[scale=0.26]{./pics/Torus.png}}
	\caption{\label{fig:Torus} The torus has one connected component($\beta_0$), two circular holes ($\beta_1$,the one in the center and the one in the middle of the "donut"), and one two-dimensional void ($\beta_2$, the inside of the "donut"). Copied from \cite{betti wiki}}
\end{figure}

@@

\indent \indent Traditionally, sampling based algorithms will build a graph for configuration space. Some of these graphs, such as \emph{PRM*} and \emph{RRT*} \cite{prmstar}, are proved to be able to find near optimal path, but can hardly be used to analysis the topology of paths. We want to to build a stronger representation of \emph{C-space} that is both able to cover optimal path and study the topological structure. In this section, we are trying to solve the problem in 2D and answer the question "Is two paths in the same homotopy class" as a start point of using our method to study the topology of \emph{C-space}. In this section, we will first give the outline of the algorithm, and explain each step in detail.

\indent In the rest of the section, we will use the term "configuration" to refer to a point in configurations space.  Ball $B(x, \rho(x))$ is referred to as a $"sample"$ in the rest of the section.

%====================================
% Algorithm outline
%====================================
\paragraph{A} \emph{The Algorithm outline} \hfill \\

@@

%%%%%%%%%%%%%%%%================================================>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
%%%%%%  We need to name our algorithm
%%%%%%  Seriously
%---------------------------------------------------------------------------------------------
\begin{figure}
  \begin{algorithmic}[1]
  \Function{RichCspaceRepresent}{\emph{C-space}, number}

@@

  \EndFunction
  \end{algorithmic}
  \caption{\label{fig:Algo} Algorithm Outline. \emph{C-space} here means the metric space $(X, \rho)$ where we can know if a configuration is feasible and its clearance to nearest obstacles.}
\end{figure}

%====================================
% Covering C-space
%====================================
\paragraph{B} \emph{Sampling in C-space} \hfill \\
\indent As pointed out in Figure \ref{fig:Algo}, this phase has two parts: Sampling on medial axis, and sampling the rest of free space.

@@

\indent Getting configurations inside a ball doesn't have to be $O(n)$ costly. It can be done using KNN algorithm, which is more efficient than iterating everything. Since we generated random configurations $n$ times(not all are feasible), in a 2D space with area $A$. We expect to have $n/A$ configurations in an unit area. A ball with radius $r$ has area $\pi r^2$, approximately
\begin{equation}
k = n \cdot \pi r^2 / A
\end{equation} 
configurations are inside the ball. We then project these (uniformly distributed) configurations to the boundary of the ball to sample new spheres. Actually, choosing a larger $k$ could be better: we virtually extend the radius of our discs, which results in a smaller density in arcs that are very close to obstacles than those far from obstacles. Samples got in this part are called \emph{regular samples}.

@@

\indent In the first part, constructing the dual shape using only the medial axis sample could break the topology of medial axis. The reason is shown in Figure \ref{badthing}: three balls are all centered on medial axis, but their dual shape will be an empty triangle, which introduces a "hole" in that area. To solve this problem we can try to sample a ball centered at the radical center after getting an empty triangle, and rebuild the dual shape for these four balls\cite{radical center}. A simple proof: if there is no obstacles in that area, the newly sampled ball will connect three balls, if there is obstacle, the ball will either not be sampled or not connect three of the balls.

@@

\indent The second part of this phase can be stated as: retract all connected 2-simplices into one point, reconnect these retracted points with other ones using their previous connection. By doing so, we got a simple graph that characterize the topology of \emph{C-space}. We would like to call it \emph{Topology Roadmap}. This roadmap can also be used to determine if paths exist between two configurations. An example of such roadmap is shown in Figure \ref{simplifiedgraph}. Filled triangles, along with large balls, are suggesting some large open areas, which is a problem \cite{largeconvex} is trying to address.

\indent This is not the only method of building the topology roadmap of \emph{C-space}, another option will be to build a visibility roadmap, like \cite{vprm}, but using only ball centers as roadmap nodes. One might come up with other heuristics to achieve the same thing. Due to the limited number of balls, any algorithms can have a good performance.

@@

\begin{figure}
	\center{\includegraphics[scale=0.3]{./pics/my_dualshape.PNG}}
	\caption{\label{ourdualshape}} Dual shape of balls generated in phase 1 (centered on medial axis). Filled triangles, along with large balls, are suggesting some large open areas.  
\end{figure}

@@

\indent Determine if a simple graph has a loop is very easy using the first Betti number $\beta_1$ mentioned in section 2, equation (1).

@@

\indent By breaking a graph into several parts, we virtually break the space into several subspaces. In this sense, our method can be viewed as a space partition strategy.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Properties of Our Algorithm
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Good Properties}\label{properties}
\indent\indent This section should be re-organized as below:
  \subsection{Coverage Analysis}
  \begin{theorem}
  Let $S_{init}$ be the set of initial random configurations. With minimum radius of $r_{min}$, configurations with clearance larger than $r_{min}$ will be covered, when $|S_{init}| \rightarrow \infty$.
  \end{theorem}

  \paragraph{\emph{Proof:}} 
  \indent When the number of initial random configurations approaches infinite, the boundaries of all existing balls are fully covered. So, every configuration not inside any existing balls will be sampled as a ball, unless it's less than $r_{min}$ clearance.\\

  As we can see, we can analyze the coverage of the space. Even more, we can explicitly describe the coverage of each ball in the algorithm. Using this property, we can find near optimal path. We will show some experiments in section 5.

@@

  \subsection{Low Memory Representation}  
  \begin{theorem}
  Let $B$ be the set of ball samples got in phase one. Then $|B|$ is finite.   
  \end{theorem}
  \paragraph{\emph{Proof:}} 
  \indent We restrict balls to have no less than $r_{min}$ radius and new balls to be centered outside existing ones. Assume we partition the space into $r_{min} \times r_{min}$ cells in 2D. The number of cells is finite, if the space is finite. Putting balls with exactly $r_{min}$ radius in the centers of these cells will cover the space totally. The number of balls is finite. Phase one will generate balls no more than the number of balls generated in this way. Figure \ref{ourdualshape} shows the dual shape of balls generated by previous phase.  \\

  In practice, most balls are much larger than minimum radius. So the actually number of balls is very limited. We will see, in section 5, the number of balls required to cover a space tends to be stable.

  \subsection{Analysis of Topology}

  In this subsection, we will see how to use the structure built by our algorithm to determine path topology.

  \begin{theorem}
  Any path $\sigma$ can be continuously deformed onto medial axis. 
  \end{theorem}   
  \paragraph{\emph{Proof}}
  \indent This is the definition of Medial Axis $MA(C_{free})$ being a strong deformation retraction of $C_{free}$.\\

  Given the theorem that any path can be deformed onto medial axis, we now use medial axis samples to determine if two paths have the same homotopy type.
